Parallel Information Set Generation for Kriegspiel

### 2012-07-15 sbac/paper.tex
removed some vspaces

\newenvironment{mytinylisting}
{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize\texttt}
{\end{list}}
\newenvironment{proof}{\vspace{-0.15in}\noindent{\bf Proof:}}%
        {\hspace*{\fill}$\Box$\par}
\newenvironment{proofsketch}{\noindent{\bf Proof Sketch.}}%
        {\hspace*{\fill}$\Box$\par\vspace{4mm}}
\newenvironment{proofof}[1]{\smallskip\noindent{\bf Proof of #1.}}%
        {\hspace*{\fill}$\Box$\par}

@@

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\vspace{-0.2in}
\end{figure}

%\begin{figure}
%\begin{verbatim}
%1. d4 {(:)}
%   a5 {(:)}
%2. Bg5 {(:)}
%   b6 {(:)}
%3. Nc3 {(:Bd8)}
%   c5 {(:)}
%4. d5 {(:Bd8;Tc5)}
%   Na6 {(:)}
%5. d6 {(:)}
%   f6 {(:Td6)}
%6. e4 {(:)}
%   fxg5 {(Xg5:Td6;Tg5)}
%7. Bc4 {(:)}
%   exd6 {(Xd6:Td6)}
%8. Qd5 {(:)}
%   Qc7 {(:)}
%9. Qf7+ {(:)} 
%   d8 {:g6,f7,e7}
%10. Qxf8++ {(1-0:)}
%\end{verbatim}
%\label{listing}
%\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.}
%\end{figure}

\section{Information Set Generation}
\label{info}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{minipage}[b]{0.35\linewidth}
\centering
\small
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
\end{verbatim}
\end{minipage}
\begin{minipage}[b]{0.25\linewidth}
\small
\centering
\begin{verbatim}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\end{minipage}
\vspace{-0.1in}
\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\vspace{-0.2in}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
information set for black after both player's have made five moves.  From this list of possibilities, black can infer
that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
black's observations up to that point.

@@

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\vspace{-0.2in}
\caption{Abbreviated output from example game.  Generating information sets for black after five turns by each player.
From this, black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\vspace{-0.3in}
\end{figure}

@@

\textbf{The Information Set Generation Algorithm}
(ISG) can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of observations for one player in a Kriegspiel game.

The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
legal moves.  Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
found to be inconsistent with the actual observation given in the input, the
subtree rooted at that node is pruned (i.e., search is discontinued along that
path).  There are types of observations that must match in order for the state
to be found: check status, pawn tries, and illegal moves available/attempted.
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\label{codelisting1}
\end{algorithmic}
\end{boxedminipage}
\caption{Algorithm}
\end{figure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\vspace{-0.1in}

@@

\label{speedups}
\vspace{-0.3in}
\end{table}

@@

In our next round of experiments, we set out to compare load-balancing
strategies for the problem.  We used the random, work stealing, and
neighborhood balancing strategies that are available in the general purpose
search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
show that there was actually little variation in performance across the
different strategies.  Across all of our experiments, the work stealing
approach appeared to be consistently---if only marginally---better than the
others.  Our hypothesis here is that the nodes are rarely starved for work.  In
other words, perhaps the very nature of the problem means that the load is
naturally well-balanced and that the load-balancer rarely needs to do a
significant amount of redistribution.  This is probably also related to the
fact that we are doing an all-solutions search. 
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
\vspace{-0.1in}

\caption{(Top) Log-log plot of running time for three different problem sizes
as the number of cores used increases.  (Bottom) Same data with a linear y-axis
for the running time (to help see the isoefficiency.)}

\caption{Effect of load balancing strategy}
\label{3schemes}
\vspace{-0.2in}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\vspace{-0.1in}
\caption{Grainsize Control for different problem sizes}
%\vspace{-0.1in}
\label{fig:dep}
\vspace{-0.1in}
\end{figure*}

@@

\begin{figure}[ht]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Kriegspiel position from 2 games, position 10}
\label{prob4} 
\vspace{-0.2in}
\end{figure}

@@

\vspace{-0.2in}
\label{tab:prob4}
\end{table}

@@

\caption{Running times, in seconds and speedups (after $|$), to find full
information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
Optimal grainsize for each number of cores is shown in bold.}

\vspace{-0.2in}
\label{tab:prob5}
\end{table}

We also evaluated our information set generation routine on 40 positions from
actual kriespiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth.

The running times and speedups for two of the more challenging instances on
Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.   These tables
show the results for runs on one to $1,024$ cores.  Each column corresponds to
a different grainsize parameter.  In this case, the grainsize is the number of
levels of the search tree that are searched in parallel.  Up to that threshold,
each node that is generated spawns a new piece of work, which the {\sc Charm++}
runtime schedules to execute on one of the processing elements according to its
load balancing strategy.  Above the threshold the entire subtree is searched to
completion on a single processor, with no additional parallel overhead.

@@

Note that both problems scale well even up to 1,024 processors.  We achieve a
speedup of $443$ in the game between humans and up to $536$ in the game between
computers.  Figures~\ref{prob4} and~\ref{prob5} show the plots of the running
times for the best grainsize for each core setting.  These plots show that the
running time continues to decrease as the number of cores increases.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\vspace{-0.2in}
\label{bestspeedups}
\end{table}

>>>>

\newenvironment{mytinylisting}
{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\small\texttt}
{\end{list}}

\newenvironment{proof}{\vspace{-0.15in}\noindent{\bf Proof:}}%
        {\hspace*{\fill}$\Box$\par}
\newenvironment{proofsketch}{\noindent{\bf Proof Sketch.}}%
        {\hspace*{\fill}$\Box$\par\vspace{4mm}}
\newenvironment{proofof}[1]{\smallskip\noindent{\bf Proof of #1.}}%
        {\hspace*{\fill}$\Box$\par}

@@

\begin{figure}
\begin{center}
\begin{verbatim}
1. d4 {(:)}
   a5 {(:)}
2. Bg5 {(:)}
   b6 {(:)}
3. Nc3 {(:Bd8)}
   c5 {(:)}
4. d5 {(:Bd8;Tc5)}
   Na6 {(:)}
5. d6 {(:)}
   f6 {(:Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   exd6 {(Xd6:Td6)}
8. Qd5 {(:)}
   Qc7 {(:)}
9. Qf7+ {(:)} 
   Kd8 {:g6,Kxf7,Ke7}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\end{center}
\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves. The final, legal move for each player is shown followed by the list of moderator declarations and all moves that were attempted but disallowed before the legal move was accepted.}
\label{listing}
\end{figure}

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

\section{Information Set Generation}
\label{info}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figures ~\ref{listing} show a full game listing and the associated pictorial progression of an actual
Kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel~\cite{wolfe07exploiting}.  Each numbered entry shows one complete turn for both White and Black.  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn try announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.

Figure~\ref{fig:figure1} shows the true state of the game after nine moves by
each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
that poin. At move 9, Black attempts to block a
potential threat along the diagonal by moving a pawn to g6, then attempts to
escape or capture a threat at f7.  When both of these attempts fail, Black can
infer that his king is being threatened by a protected bishop or queen at f7.
The attempt to move to e7 is Black's way of finding out whether the threatening
piece is a queen or bishop.  When this move fails, black knows it is White's
queen and ultimately retreats to d8.  Note that the failed moves by White on
turns 3 and 4 are a result of being blocked by the pawn at e2.

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from White's perspective is shown in Figure~\ref{filteredlisting}.  Note that Black's actual moves have been
replaced with ``??'' and lists of Black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, Black attempted three illegal moves.
\begin{figure}
%\begin{minipage}[b]{0.35\linewidth}
%\centering
%\small
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
%\end{verbatim}
%\end{minipage}
%\begin{minipage}[b]{0.25\linewidth}
%\small
%\centering
%\begin{verbatim}
%\end{minipage}
%\vspace{-0.1in}
\caption{Observations received by White over the course of the game.  White does not know Black's actual moves but hears moderator announcements related to captures, pawn tries, and the number of attempted moves by Black that fail.}
\label{filteredlisting}
%\vspace{-0.2in}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for Black is in Black's information set.  Figure~\ref{abbrevoutput} shows a portion of the
information set for Black after both player's have made five moves.  From this list of possibilities, Black can infer
that White must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
Black's observations up to that point.

@@

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\caption{Abbreviated output from example game.  Three of 25 sequences that are consistent with Black's observations after five moves by each player.  The 25 sequences (or the states that result from them) compose Black's information set at that point in the game. From this, Black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\end{figure}

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.

The nodes in the search tree at depth $n$ correspond possible to possible move 
sequences of $n$ plies from the start state.  (A sequence of $t$ actions from
the start state is denoted $a_{1:t}$.) There is a search space variable for
each ply in the game, and the values that a variable can take on are all
possible legal moves.  Information set generation, then, is an all-solutions
search through the space.  Whenever a node in the search tree is found to be
inconsistent with the actual observation given in the input, the subtree rooted
at that node is pruned (i.e., search is discontinued along that path).  There
are three types of observations that must match in order for a state $s$ to be
included in the information set $I$ at depth $t$: check status, pawn tries, and
illegal moves available/attempted.

\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Algorithm}
\label{codelisting1}
\end{figure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}

@@

\label{speedups}
\end{table}

@@

In our next round of experiments, we set out to compare load-balancing
strategies for the problem.  We used the random, work stealing, and
neighborhood balancing strategies that are available in the general purpose
search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
show that there was actually little variation in performance across the
different strategies.  Across all of our experiments, the work stealing
approach appeared to be consistently---if only marginally---better than the
others.  Our hypothesis here is that the nodes are rarely starved for work.  In
other words, perhaps the very nature of the problem means that the load is
naturally well-balanced and that the load-balancer rarely needs to do a
significant amount of redistribution.  This is probably also related to the
fact that we are doing an all-solutions search. 
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}

\caption{Effect of load balancing strategy}
\label{3schemes}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\caption{Grainsize Control for different problem sizes}
%\vspace{-0.1in}
\label{fig:dep}
\end{figure*}
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time, to help see the isoefficiency.}
\label{3depths}
\end{figure}

@@

\begin{figure}[ht]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Kriegspiel position from 2 games, position 10}
\label{prob4} 
\end{figure}

@@

\label{tab:prob4}
\end{table}

@@

\caption{Running times, in seconds and speedups (after $|$), to find full
information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
Optimal grainsize for each number of cores is shown in bold.}
\label{tab:prob5}
\end{table}

We also evaluated our information set generation routine on 40 positions from
actual Kriespiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth.

The running times and speedups for two of the more challenging instances on
Ranger are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables
show the results for runs on one to $1,024$ cores.  Each column corresponds to
a different grainsize parameter.  In this case, the grainsize is the number of
levels of the search tree that are searched in parallel.  Up to that threshold,
each node that is generated spawns a new piece of work, which the {\sc Charm++}
runtime schedules to execute on one of the processing elements according to its
load balancing strategy.  Above the threshold the entire subtree is searched to
completion on a single processor, with no additional parallel overhead.

@@

Note that both problems scale well even up to 1,024 processors.  We achieve a
speedup of $443$ in the game between humans and up to $536$ in the game between
computers.  Figure~\ref{prob4} shows the plots of the running
times on these two problems with the best grainsize for each core setting.  These plots show that the
running time continues to decrease as the number of cores increases.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\label{bestspeedups}
\end{table}
